\documentclass[xcolor=table,dvipsnames]{beamer}

\usepackage{lscape, amsmath, amsfonts, amssymb, setspace, theorem, wrapfig, graphicx, float, multirow, subfig, color, rotating, multicol, datetime, natbib, venndiagram, pstricks, xkeyval, tikz, etoolbox, verbatim, pgf, tikz, pgfplots, mathrsfs, nth}

\usepackage{listings}
\usepackage{xcolor}
\usetikzlibrary{arrows}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegreengray}{rgb}{0,0.4,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codeblue}{rgb}{0.00,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\definecolor{jeopardy}{rgb}{.24,.47,.914}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreengray},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codegreen},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
 
\lstset{style=mystyle}

\title{1X: Introduction to R}
\subtitle{Essex Summer School in Social Science Data Analysis -- University of Essex}
\date{11 July, 2021}
\author{Lorenzo Crippa}
\institute{University of Essex -- Department of Government}

\usetheme[progressbar=frametitle]{metropolis}
\usecolortheme{seahorse}

\begin{document}

\begin{frame}[plain]
\begin{center}
\titlepage
\end{center}
\end{frame}

\begin{frame}{Today's session}
Welcome! My name is Lorenzo, I am going to be your instructor for today. We also have Phil today, who will help us.

Today we'll introduce R. \\ \pause
The goals of today's session are to: \pause
\begin{enumerate}
\item[0.] Introduce what R and RStudio are \pause
\item Learn about data types, R objects, and how to create them \pause
\item Learn how to import and manage datasets \pause
\item Learn how to generate basic plots and summary statistics \pause
\item Learn how to run basic statistical analyses
\end{enumerate}
\end{frame}

\begin{frame}{Session structure}
Today's class is structured in two main sessions: \pause
\begin{enumerate}
\item Data types, R objects, data manipulation (3h) \pause
	\begin{enumerate}
	\item Hands-on explanation
	\item Exercises
	\end{enumerate} \pause
\item Plotting and basic summary statistics (2:30h) \pause
	\begin{enumerate}
	\item Hands-on explanation
	\item Exercises
	\end{enumerate}
\end{enumerate} \pause

A break between sessions one and two \pause

!! \textbf{Always} feel free to interrupt me during the presentation for questions or clarifications by raising your virtual hand. I'll try to check on them as often as I can. \pause Worst case scenario: feel free to interrupt me while I speak.
\end{frame}

\section{0. Introduction to R and RStudio}

\begin{frame}{What is R?}
\begin{itemize}
\item A \textbf{programming language} developed specifically for \textbf{professional} data analysis (based on S). \pause No point and click \pause
\item It is a powerful language. It is suited to two types of works: \pause
	\begin{enumerate}
	\item An interactive work; users employ packages and functions that were already written (today's session) \pause
	\item A programming work, more advanced; users write new functions and contribute to the language \pause
	\end{enumerate}
%\item R has a rather steep learning curve, but the passage from 1. to 2. comes quicker than expected \pause
\item R is free and open source, and there is an incredible community that uses the language and contributes to it \pause
\item We need to have R installed on our computer to make it run
\end{itemize}
\end{frame}

\begin{frame}{What is RStudio?}
We run R in RStudio. What is RStudio? \pause
\begin{itemize}
\item It is an integrated development environment (IDE) \pause
\item It automatically flags syntax errors and assigns different colors to different chunks of code \pause
\item The basic version (the one you'll need 99.999\% of the times) is free \pause
\item We still need to have R installed on our computer in order for RStudio to properly run
\end{itemize}

\end{frame}

\section{1. Data types, R objects, data manipulation}

\section{2. Plotting and basic summary statistics}

\section{Extra topics}

\begin{frame}{Useful R packages}
Some R packages we haven't covered but that might be useful to some of you: \\ \pause

\begin{center}
\resizebox{10cm}{!}{
\begin{tabular}{lll}
\textbf{Type} 		& \textbf{Package name} 	& \textbf{Aim} \\
Data import			& \texttt{foreign} 	& Imports \texttt{.dta} files when \texttt{haven} doesn't \\
					& \texttt{readr}	& Imports files from various formats \\
\hline
Data cleaning 		& \texttt{reshape} 	& Turning long datasets into wide and viceversa\\
					& \texttt{tidyr} 	& Tidy datasets \\
\hline
Plots 				& \texttt{ggplot2} 	& Plots based on Grammar of Graphics \\
					& \texttt{lattice}	& Plots, less versatile than \texttt{ggplot2} \\
\hline
Standard errors		& \texttt{sandwich} + \texttt{lmtest}  & Robust and clustered standard errors + test coefficients \\
					& \texttt{estimatr} & Alternative to \texttt{sandwich} and \texttt{lmtest} \\
\hline
Modelling			& \texttt{plm} 		& Panel data models (within-between estimator) \\
					& \texttt{ggeffects}& Marginal effects from models \\
					& \texttt{AER} 		& Various (Instrumental variable, tobit models) \\ 
					& \texttt{MASS} 	& Various (ordered logit and probit) \\ 
					& \texttt{mlogit} or \texttt{mnlogit}	& Multinomial logit models \\
					& \texttt{lme4} 	& Multilevel models \\
					& \texttt{rdd} 		& Regression discontinuity design \\
					& \texttt{zoo} or \texttt{forecast} & Time series tools \\
\end{tabular}
}
\end{center}
\end{frame}

\begin{frame}{That's all}
\centering
Thank you for the attention! \smallskip

Lorenzo Crippa\\
l.crippa@essex.ac.uk
\end{frame}
\end{document}